Evaluating FAIR Digital Object and Linked Data as distributed object systems

### 2023-06-19 latex/table1.tex

  & HTTP/1.1 TCP/IP port 80 \autocite{rfc2616}, HTTP/1.1+TLS TCP/IP 443 \autocite{rfc2818}, HTTP/2 (as HTTP/1* but binary) \autocite{rfc7540}, HTTP/3 (like HTTP/2+TLS but UDP) \autocite{rfc9114} \\

@@

  & Main DOIP implementation is \href{https://www.cordra.org/}{Cordra}, which can be single-instance or \href{https://www.cordra.org/documentation/configuration/distributed-deployment.html}{distributed}. Cordra \href{https://www.cordra.org/documentation/configuration/storage-backends.html}{storage backends} include file system, S3, MongoDB (itself scalable). Unique DOIP protocol can be hard to add to existing Web application frameworks, although proxy services have been developed (e.g.~B2SHARE adapter).
  & HTTP services widely deployed in a myriad of ways, ranging from single instance servers, horizontally \& vertically scaled application servers, to (for static content) multi-cloud Content-Delivery Networks (CDN). Current scalable cloud technologies for Web hosting may not support HTTP features previously seen as important for Semantic Web, e.g.~content negotiation and semantic HTTP status codes. \\

>>>>

  & HTTP/1.1, TCP/IP port 80 \autocite{rfc2616}; HTTP/1.1+TLS, TCP/IP 443 \autocite{rfc2818}; HTTP/2, as HTTP/1* but binary \autocite{rfc7540}; HTTP/3, like HTTP/2+TLS but UDP \autocite{rfc9114} \\

@@

  & Main DOIP implementation is \href{https://www.cordra.org/}{\emph{Cordra}}, which can be single-instance or \href{https://www.cordra.org/documentation/configuration/distributed-deployment.html}{distributed}. Cordra \href{https://www.cordra.org/documentation/configuration/storage-backends.html}{storage backends} include file system, S3, MongoDB (itself scalable). Unique DOIP protocol can be hard to add to existing Web application frameworks, although proxy services have been developed (e.g.~B2SHARE adapter).
  & HTTP services widely deployed in a myriad of ways, ranging from single instance servers, horizontally \& vertically scaled application servers, to multi-cloud Content-Delivery Networks (CDN). Current scalable cloud technologies for Web hosting may not support HTTP features previously seen as important for Semantic Web, e.g.~content negotiation and semantic HTTP status codes. \\

### 2023-06-19 latex/table3.tex
more bib tweaks.

@@

  & HTTP stable for 30 years, Semantic Web for 20 years. \texttt{http://} URIs replaced by \texttt{https://}.

@@

  & Add explicit expression to show each FAIR pcinciple fulfilled. \\

@@

  & Allow method discovery by type FDOs in advance (see PR-TypingFDOs-2.0-20220608).

@@

  & HTTP, LDP. FDOF not yet standardised
  & Formalise FDOF from FDOF-SEM working group \\

@@

  & Clarify how local testing handles can be used during development, how ``temporary'' FDOs should evolve \autocite{fdo-PIDProfileAttributes}. Register \texttt{0.DOIP/*} and \texttt{0.FDO/*} as PIDs.

@@

  & Add strong guidance for PID services like w3id and persistence policies. \\

@@

  & Add FAIR Signposting at PID provider for minimal PID record \\

@@

  & Freetext attribute keys. Attributes should be defined for FDO type (?).

@@

  & Defines open-ended ``Response Attributes'' without namespaces, but mandated as ``None'' for all CRUD operations. Metadata would need to be bundled within custom FDO types or attributes. Unclear how levels are separated within single FDO representation (need FDOF8?).

@@

  & Metadata schemas are in practice managed on single CORDA server as local types, using JSON Schema.

@@

>>>>

\newgeometry{left=4cm,%
             right=2cm,%
             top=2.25cm,%
             bottom=2.25cm,%
             headheight=12pt}

@@

  & HTTP stable for 30 years, Semantic Web for 20 years. \texttt{http://} URIs mostly replaced by \texttt{https://}.

@@

  & Add explicit expression to show each FAIR principle fulfilled. \\

@@

  & Allow method discovery by type FDOs in advance \autocite{fdo-TypingFDOs}.

@@

  & HTTP, LDP. However FDOF is not yet standardised.
  & Formalise FDOF from FDOF-SEM working group. \\

@@

  & Clarify how local testing handles can be used during development, how ``temporary'' FDOs should evolve \autocite{fdo-PIDProfileAttributes}. Register \texttt{0.DOIP/*} and \texttt{0.FDO/*} as actual PIDs.

@@

  & Add strong guidance for PID services like w3id and persistence policies \autocite{McMurry_2017}. \\

@@

  & Add FAIR Signposting \autocite{vandesompelFAIRSignpostingProfile2022} at PID provider for minimal PID record \\

@@

  & Freetext attribute keys. Attributes should be defined for FDO type.

@@

  & Defines open-ended ``Response Attributes'' without namespaces, but mandated as ``None'' for all CRUD operations. Metadata would need to be bundled within custom FDO types or attributes. Unclear how levels are separated within single FDO representation (may need FDOF8).

@@

  & Metadata schemas are in practice managed on single Cordra server as local types, using JSON Schema.

@@

\restoregeometry

### 2023-06-19 latex/table4.tex

	\caption[Comparing FAIR Digital Object and Web technologies as middleware infrastructures]{Comparing FAIR Digital Object (with the DOIP 2.0 protocol \cite{13TcbsZF6}) and Web technologies (using Linked Data) as middleware infrastructures \cite{hRzcHhPD}

>>>>

	\caption[Comparing FAIR Digital Object and Web technologies as middleware infrastructures]{Comparing FAIR Digital Object (with the DOIP 2.0 protocol \cite{foundationDigitalObjectInterface}) and Web technologies (using Linked Data) as middleware infrastructures \cite{zarrasComparisonFrameworkMiddleware2004a}

### 2023-06-19 latex/table5.tex

  & Not specified

@@

  & & Ubiquitous, hundreds of HTTP libraries \\

@@

    One DOIPv2 implementation (\href{https://www.cordra.org/}{CORDRA}): free under BSD-like license (not recognised as Open Source).    

@@

  & Unspecified (some CORDRA types add getProvenance methods). PID Kernel attributes? Unspecified W3C PROV-O, PAV
  & & \\

>>>>

@@

  & HTTP \texttt{GET}, content-negotiation for \texttt{fdof/object}
  & Ubiquitous, hundreds of HTTP libraries \\

@@

    One DOIPv2 implementation (\href{https://www.cordra.org/}{Cordra}): free under BSD-like license (not recognised as Open Source).    

@@

  & Unspecified (some Cordra types add getProvenance methods). PID Kernel attributes? 
  & --- & W3C PROV-O, PAV \\
